---
title: "How do shooting incident breakdowns by race and gender in NYC's boroughs compare to the city overall?"
author: Jacob Moore
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data Source
[Webpage](https://catalog.data.gov/dataset/nypd-shooting-incident-data-historic/resource/c564b578-fd8a-4005-8365-34150d306cc4)

[CSV](https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD)

```{r define_shooting_data, echo=FALSE}
shooting_data <- read.csv("https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD")
num_shootings_all <- nrow(shooting_data)
```

```{r clean_data, echo=FALSE}

clean_data <- subset(shooting_data, select = -c(INCIDENT_KEY, OCCUR_DATE, OCCUR_TIME, PERP_AGE_GROUP, JURISDICTION_CODE, LOCATION_DESC, PRECINCT, STATISTICAL_MURDER_FLAG, X_COORD_CD, Y_COORD_CD, Latitude, Longitude, Lon_Lat))


```

### NYC Overall Data Summary
```{r data_summary, echo=FALSE}
num_shootings <- nrow(clean_data)

female_victims <- subset(clean_data, VIC_SEX=='F')
male_victims <- subset(clean_data, VIC_SEX=='M')
num_female_victims <- nrow(female_victims)
num_male_victims <- nrow(male_victims)
percent_female <- round((num_female_victims / num_shootings) * 100, digits=2)
percent_male <- round((num_male_victims / num_shootings) * 100, digits=2)

white_victims <- subset(clean_data, VIC_RACE=='WHITE')
num_white_victims <- nrow(white_victims)
percent_white <- round((num_white_victims / num_shootings) * 100, digits=2)

black_victims <- subset(clean_data, (VIC_RACE=='BLACK' | VIC_RACE=='BLACK HISPANIC'))
num_black_victims <- nrow(black_victims)
percent_black <- round((num_black_victims / num_shootings) * 100, digits=2)

hispanic_victims <- subset(clean_data, (VIC_RACE=='WHITE HISPANIC' | VIC_RACE=='BLACK HISPANIC'))
num_hispanic_victims <- nrow(hispanic_victims)
percent_hispanic <- round((num_hispanic_victims / num_shootings) * 100, digits=2)

other_victims <- subset(clean_data, (VIC_RACE!='WHITE' & VIC_RACE!='BLACK' & VIC_RACE!='WHITE HISPANIC' & VIC_RACE!='BLACK HISPANIC'))
num_other_victims <- nrow(other_victims)
percent_other <- round((num_other_victims / num_shootings) * 100, digits=2)
```

There were `r num_shootings_all` shooting incidents recorded by the NYPD in this data set in the city overall.

The victims of these shootings were `r percent_female`% female, and `r percent_male`% male.

`r percent_black`% of victims were Black and `r percent_hispanic`% were Hispanic, with only `r percent_white`% being White and non-Hispanic. `r percent_other`% of shooting victims belonged to other races or were listed as being of unknown race. Note that percentages sum to more than 100 because some victims are listed as both Black and Hispanic or White and Hispanic.

Victims of shootings were thus, overwhelmingly male and Black.


### Boroughs Data Summaries


```{r bronx, echo=FALSE}
my_boro <- 'BRONX'
local_data <- subset(clean_data, BORO==my_boro)

num_shootings <- nrow(local_data)

percent_of_shootings <- round((num_shootings / num_shootings_all) * 100, digits=2)

female_victims <- subset(local_data, VIC_SEX=='F')
male_victims <- subset(local_data, VIC_SEX=='M')
num_female_victims <- nrow(female_victims)
num_male_victims <- nrow(male_victims)
percent_female <- round((num_female_victims / num_shootings) * 100, digits=2)
percent_male <- round((num_male_victims / num_shootings) * 100, digits=2)

white_victims <- subset(local_data, VIC_RACE=='WHITE')
num_white_victims <- nrow(white_victims)
percent_white <- round((num_white_victims / num_shootings) * 100, digits=2)

black_victims <- subset(local_data, (VIC_RACE=='BLACK' | VIC_RACE=='BLACK HISPANIC'))
num_black_victims <- nrow(black_victims)
percent_black <- round((num_black_victims / num_shootings) * 100, digits=2)

hispanic_victims <- subset(local_data, (VIC_RACE=='WHITE HISPANIC' | VIC_RACE=='BLACK HISPANIC'))
num_hispanic_victims <- nrow(hispanic_victims)
percent_hispanic <- round((num_hispanic_victims / num_shootings) * 100, digits=2)

other_victims <- subset(local_data, (VIC_RACE!='WHITE' & VIC_RACE!='BLACK' & VIC_RACE!='WHITE HISPANIC' & VIC_RACE!='BLACK HISPANIC'))
num_other_victims <- nrow(other_victims)
percent_other <- round((num_other_victims / num_shootings) * 100, digits=2)
```

#### Bronx Data Summary

There were `r num_shootings` shooting incidents recorded by the NYPD in this data set in the Bronx, `r percent_of_shootings`% of the total shooting incidents in New York City.

The victims of these shootings were `r percent_female`% female, and `r percent_male`% male.

`r percent_black`% of victims were Black and `r percent_hispanic`% were Hispanic, with only `r percent_white`% being White and non-Hispanic. `r percent_other`% of shooting victims belonged to other races or were listed as being of unknown race. Note that percentages sum to more than 100 because some victims are listed as both Black and Hispanic or White and Hispanic.



```{r brooklyn, echo=FALSE}
my_boro <- 'BROOKLYN'
local_data <- subset(clean_data, BORO==my_boro)

num_shootings <- nrow(local_data)

percent_of_shootings <- round((num_shootings / num_shootings_all) * 100, digits=2)

female_victims <- subset(local_data, VIC_SEX=='F')
male_victims <- subset(local_data, VIC_SEX=='M')
num_female_victims <- nrow(female_victims)
num_male_victims <- nrow(male_victims)
percent_female <- round((num_female_victims / num_shootings) * 100, digits=2)
percent_male <- round((num_male_victims / num_shootings) * 100, digits=2)

white_victims <- subset(local_data, VIC_RACE=='WHITE')
num_white_victims <- nrow(white_victims)
percent_white <- round((num_white_victims / num_shootings) * 100, digits=2)

black_victims <- subset(local_data, (VIC_RACE=='BLACK' | VIC_RACE=='BLACK HISPANIC'))
num_black_victims <- nrow(black_victims)
percent_black <- round((num_black_victims / num_shootings) * 100, digits=2)

hispanic_victims <- subset(local_data, (VIC_RACE=='WHITE HISPANIC' | VIC_RACE=='BLACK HISPANIC'))
num_hispanic_victims <- nrow(hispanic_victims)
percent_hispanic <- round((num_hispanic_victims / num_shootings) * 100, digits=2)

other_victims <- subset(local_data, (VIC_RACE!='WHITE' & VIC_RACE!='BLACK' & VIC_RACE!='WHITE HISPANIC' & VIC_RACE!='BLACK HISPANIC'))
num_other_victims <- nrow(other_victims)
percent_other <- round((num_other_victims / num_shootings) * 100, digits=2)

```

#### Brooklyn Data Summary

There were `r num_shootings` shooting incidents recorded by the NYPD in this data set in Brooklyn, `r percent_of_shootings`% of the total shooting incidents in New York City.

The victims of these shootings were `r percent_female`% female, and `r percent_male`% male.

`r percent_black`% of victims were Black and `r percent_hispanic`% were Hispanic, with only `r percent_white`% being White and non-Hispanic. `r percent_other`% of shooting victims belonged to other races or were listed as being of unknown race. Note that percentages sum to more than 100 because some victims are listed as both Black and Hispanic or White and Hispanic.

```{r manhattan, echo=FALSE}
my_boro <- 'MANHATTAN'
local_data <- subset(clean_data, BORO==my_boro)

num_shootings <- nrow(local_data)

percent_of_shootings <- round((num_shootings / num_shootings_all) * 100, digits=2)

female_victims <- subset(local_data, VIC_SEX=='F')
male_victims <- subset(local_data, VIC_SEX=='M')
num_female_victims <- nrow(female_victims)
num_male_victims <- nrow(male_victims)
percent_female <- round((num_female_victims / num_shootings) * 100, digits=2)
percent_male <- round((num_male_victims / num_shootings) * 100, digits=2)

white_victims <- subset(local_data, VIC_RACE=='WHITE')
num_white_victims <- nrow(white_victims)
percent_white <- round((num_white_victims / num_shootings) * 100, digits=2)

black_victims <- subset(local_data, (VIC_RACE=='BLACK' | VIC_RACE=='BLACK HISPANIC'))
num_black_victims <- nrow(black_victims)
percent_black <- round((num_black_victims / num_shootings) * 100, digits=2)

hispanic_victims <- subset(local_data, (VIC_RACE=='WHITE HISPANIC' | VIC_RACE=='BLACK HISPANIC'))
num_hispanic_victims <- nrow(hispanic_victims)
percent_hispanic <- round((num_hispanic_victims / num_shootings) * 100, digits=2)

other_victims <- subset(local_data, (VIC_RACE!='WHITE' & VIC_RACE!='BLACK' & VIC_RACE!='WHITE HISPANIC' & VIC_RACE!='BLACK HISPANIC'))
num_other_victims <- nrow(other_victims)
percent_other <- round((num_other_victims / num_shootings) * 100, digits=2)

```

#### Manhattan Data Summary

There were `r num_shootings` shooting incidents recorded by the NYPD in this data set in Manhattan, `r percent_of_shootings`% of the total shooting incidents in New York City.

The victims of these shootings were `r percent_female`% female, and `r percent_male`% male.

`r percent_black`% of victims were Black and `r percent_hispanic`% were Hispanic, with only `r percent_white`% being White and non-Hispanic. `r percent_other`% of shooting victims belonged to other races or were listed as being of unknown race. Note that percentages sum to more than 100 because some victims are listed as both Black and Hispanic or White and Hispanic.


```{r queens, echo=FALSE}
my_boro <- 'QUEENS'
local_data <- subset(clean_data, BORO==my_boro)

num_shootings <- nrow(local_data)

percent_of_shootings <- round((num_shootings / num_shootings_all) * 100, digits=2)

female_victims <- subset(local_data, VIC_SEX=='F')
male_victims <- subset(local_data, VIC_SEX=='M')
num_female_victims <- nrow(female_victims)
num_male_victims <- nrow(male_victims)
percent_female <- round((num_female_victims / num_shootings) * 100, digits=2)
percent_male <- round((num_male_victims / num_shootings) * 100, digits=2)

white_victims <- subset(local_data, VIC_RACE=='WHITE')
num_white_victims <- nrow(white_victims)
percent_white <- round((num_white_victims / num_shootings) * 100, digits=2)

black_victims <- subset(local_data, (VIC_RACE=='BLACK' | VIC_RACE=='BLACK HISPANIC'))
num_black_victims <- nrow(black_victims)
percent_black <- round((num_black_victims / num_shootings) * 100, digits=2)

hispanic_victims <- subset(local_data, (VIC_RACE=='WHITE HISPANIC' | VIC_RACE=='BLACK HISPANIC'))
num_hispanic_victims <- nrow(hispanic_victims)
percent_hispanic <- round((num_hispanic_victims / num_shootings) * 100, digits=2)

other_victims <- subset(local_data, (VIC_RACE!='WHITE' & VIC_RACE!='BLACK' & VIC_RACE!='WHITE HISPANIC' & VIC_RACE!='BLACK HISPANIC'))
num_other_victims <- nrow(other_victims)
percent_other <- round((num_other_victims / num_shootings) * 100, digits=2)
```

#### Queens Data Summary

There were `r num_shootings` shooting incidents recorded by the NYPD in this data set in the Queens, `r percent_of_shootings`% of the total shooting incidents in New York City.

The victims of these shootings were `r percent_female`% female, and `r percent_male`% male.

`r percent_black`% of victims were Black and `r percent_hispanic`% were Hispanic, with only `r percent_white`% being White and non-Hispanic. `r percent_other`% of shooting victims belonged to other races or were listed as being of unknown race. Note that percentages sum to more than 100 because some victims are listed as both Black and Hispanic or White and Hispanic.


```{r staten_island, echo=FALSE}
my_boro <- 'STATEN ISLAND'
local_data <- subset(clean_data, BORO==my_boro)

num_shootings <- nrow(local_data)

percent_of_shootings <- round((num_shootings / num_shootings_all) * 100, digits=2)

female_victims <- subset(local_data, VIC_SEX=='F')
male_victims <- subset(local_data, VIC_SEX=='M')
num_female_victims <- nrow(female_victims)
num_male_victims <- nrow(male_victims)
percent_female <- round((num_female_victims / num_shootings) * 100, digits=2)
percent_male <- round((num_male_victims / num_shootings) * 100, digits=2)

white_victims <- subset(local_data, VIC_RACE=='WHITE')
num_white_victims <- nrow(white_victims)
percent_white <- round((num_white_victims / num_shootings) * 100, digits=2)

black_victims <- subset(local_data, (VIC_RACE=='BLACK' | VIC_RACE=='BLACK HISPANIC'))
num_black_victims <- nrow(black_victims)
percent_black <- round((num_black_victims / num_shootings) * 100, digits=2)

hispanic_victims <- subset(local_data, (VIC_RACE=='WHITE HISPANIC' | VIC_RACE=='BLACK HISPANIC'))
num_hispanic_victims <- nrow(hispanic_victims)
percent_hispanic <- round((num_hispanic_victims / num_shootings) * 100, digits=2)

other_victims <- subset(local_data, (VIC_RACE!='WHITE' & VIC_RACE!='BLACK' & VIC_RACE!='WHITE HISPANIC' & VIC_RACE!='BLACK HISPANIC'))
num_other_victims <- nrow(other_victims)
percent_other <- round((num_other_victims / num_shootings) * 100, digits=2)

```

#### Staten Island Data Summary

There were `r num_shootings` shooting incidents recorded by the NYPD in this data set in Staten Island, `r percent_of_shootings`% of the total shooting incidents in New York City.

The victims of these shootings were `r percent_female`% female, and `r percent_male`% male.

`r percent_black`% of victims were Black and `r percent_hispanic`% were Hispanic, with only `r percent_white`% being White and non-Hispanic. `r percent_other`% of shooting victims belonged to other races or were listed as being of unknown race. Note that percentages sum to more than 100 because some victims are listed as both Black and Hispanic or White and Hispanic.

### Session Info
```{r session_info, echo=FALSE}
sessionInfo()
```

