---
title: "How do shooting incident breakdowns by race and gender in NYC's boroughs compare to the city overall?"
author: Jacob Moore
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data Source
[Webpage](https://catalog.data.gov/dataset/nypd-shooting-incident-data-historic/resource/c564b578-fd8a-4005-8365-34150d306cc4)

[CSV](https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD)

```{r define_shooting_data, echo=FALSE}
shooting_data <- read.csv("https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD")
num_shootings_all <- nrow(shooting_data)
```

```{r clean_data, echo=FALSE}

date_strings <- shooting_data$OCCUR_DATE

start_date <- min(date_strings)
end_date <- max(date_strings)

clean_data <- subset(shooting_data, select = -c(INCIDENT_KEY, OCCUR_DATE, OCCUR_TIME, PERP_AGE_GROUP, JURISDICTION_CODE, LOCATION_DESC, PRECINCT, STATISTICAL_MURDER_FLAG, X_COORD_CD, Y_COORD_CD, Latitude, Longitude, Lon_Lat))


```

### NYC Overall Data Summary

```{r data_summary, echo=FALSE}
num_shootings <- nrow(clean_data)

female_victims <- subset(clean_data, VIC_SEX=='F')
male_victims <- subset(clean_data, VIC_SEX=='M')
num_female_victims <- nrow(female_victims)
num_male_victims <- nrow(male_victims)
percent_female <- round((num_female_victims / num_shootings) * 100, digits=2)
percent_male <- round((num_male_victims / num_shootings) * 100, digits=2)

white_victims <- subset(clean_data, VIC_RACE=='WHITE')
num_white_victims <- nrow(white_victims)
percent_white <- round((num_white_victims / num_shootings) * 100, digits=2)

black_victims <- subset(clean_data, (VIC_RACE=='BLACK' | VIC_RACE=='BLACK HISPANIC'))
num_black_victims <- nrow(black_victims)
percent_black <- round((num_black_victims / num_shootings) * 100, digits=2)

hispanic_victims <- subset(clean_data, (VIC_RACE=='WHITE HISPANIC' | VIC_RACE=='BLACK HISPANIC'))
num_hispanic_victims <- nrow(hispanic_victims)
percent_hispanic <- round((num_hispanic_victims / num_shootings) * 100, digits=2)

other_victims <- subset(clean_data, (VIC_RACE!='WHITE' & VIC_RACE!='BLACK' & VIC_RACE!='WHITE HISPANIC' & VIC_RACE!='BLACK HISPANIC'))
num_other_victims <- nrow(other_victims)
percent_other <- round((num_other_victims / num_shootings) * 100, digits=2)
```


This data set provides a record of shooting incidents occurring in New York City between `r start_date` and `r end_date`.
There were `r num_shootings_all` shooting incidents recorded by the NYPD in this data set in the city overall.

#### Analysis

The victims of these shootings were `r percent_female`% female, and `r percent_male`% male.

```{r overall_gender_graph, echo=FALSE}
vics <- c(num_female_victims, num_male_victims)
gender_labels <- c("Female Victims", "Male Victims")
pie(vics, gender_labels, main="Proportion of Victims by Gender")
```

`r percent_black`% of victims were Black and `r percent_hispanic`% were Hispanic, with only `r percent_white`% being White and non-Hispanic. `r percent_other`% of shooting victims belonged to other races or were listed as being of unknown race. Note that percentages sum to more than 100 because some victims are listed as both Black and Hispanic or White and Hispanic.

```{r overall_race_graph, echo=FALSE}
races <- c(percent_black, percent_hispanic, percent_white, percent_other)
labels <- c("Black", "Hispanic", "White", "Other")
barplot(races, names.arg=labels, main="Shootings by Race", xlab="Race", ylab="Percentage of Shooting Victims")
```

##### Initial Analysis of Overall Data

Victims of shootings were thus, overwhelmingly male and Black, with Hispanics being the next largest percentage of victims. Very, very few shooting victims appear to be White or female.


### Boroughs Data Summaries and Analysis



#### Bronx Data Summary

```{r bronx, echo=FALSE}
my_boro <- 'BRONX'
local_data <- subset(clean_data, BORO==my_boro)

num_shootings <- nrow(local_data)


percent_of_shootings <- round((num_shootings / num_shootings_all) * 100, digits=2)

female_victims <- subset(local_data, VIC_SEX=='F')
male_victims <- subset(local_data, VIC_SEX=='M')
num_female_victims <- nrow(female_victims)
num_male_victims <- nrow(male_victims)
percent_female <- round((num_female_victims / num_shootings) * 100, digits=2)
percent_male <- round((num_male_victims / num_shootings) * 100, digits=2)

white_victims <- subset(local_data, VIC_RACE=='WHITE')
num_white_victims <- nrow(white_victims)
percent_white <- round((num_white_victims / num_shootings) * 100, digits=2)


black_victims <- subset(local_data, (VIC_RACE=='BLACK' | VIC_RACE=='BLACK HISPANIC'))
num_black_victims <- nrow(black_victims)
percent_black <- round((num_black_victims / num_shootings) * 100, digits=2)


hispanic_victims <- subset(local_data, (VIC_RACE=='WHITE HISPANIC' | VIC_RACE=='BLACK HISPANIC'))
num_hispanic_victims <- nrow(hispanic_victims)
percent_hispanic <- round((num_hispanic_victims / num_shootings) * 100, digits=2)


other_victims <- subset(local_data, (VIC_RACE!='WHITE' & VIC_RACE!='BLACK' & VIC_RACE!='WHITE HISPANIC' & VIC_RACE!='BLACK HISPANIC'))
num_other_victims <- nrow(other_victims)
percent_other <- round((num_other_victims / num_shootings) * 100, digits=2)

bronx_shootings <- num_shootings
bronx_fem <- percent_female
bronx_mal <- percent_male
bronx_white <- percent_white
bronx_black<-percent_black
bronx_hisp <- percent_hispanic
bronx_oth <- other_victims
```

There were `r num_shootings` shooting incidents recorded by the NYPD in this data set in the Bronx, `r percent_of_shootings`% of the total shooting incidents in New York City.

```{r bronx_shootings_graph, echo=FALSE}
pie(c(num_shootings, num_shootings_all-num_shootings), c("Shootings in the Bronx", "Shootings Elsewhere in NYC"), main="Proportion of Shootings in NYC that occured in the Bronx")
```


The victims of these shootings were `r percent_female`% female, and `r percent_male`% male.

```{r bronx_gender_graph, echo=FALSE}
vics <- c(num_female_victims, num_male_victims)
gender_labels <- c("Female Victims", "Male Victims")
pie(vics, gender_labels, main="Proportion of Victims by Gender")
```

`r percent_black`% of victims were Black and `r percent_hispanic`% were Hispanic, with only `r percent_white`% being White and non-Hispanic. `r percent_other`% of shooting victims belonged to other races or were listed as being of unknown race. Note that percentages sum to more than 100 because some victims are listed as both Black and Hispanic or White and Hispanic.

```{r bronx_race_graph, echo=FALSE}
races <- c(percent_black, percent_hispanic, percent_white, percent_other)
labels <- c("Black", "Hispanic", "White", "Other")
barplot(races, names.arg=labels, main="Shootings by Race", xlab="Race", ylab="Percentage of Shooting Victims")
```

##### Analysis

The Bronx is second only to Brooklyn in number of shootings. The numbers here do not diverge significantly from the numbers for the city overall in terms of gender or racial make up of shooting victims. Men and Black people still make up the vast majority of victims, though a larger share of victims here are Hispanic and Black victims make up a smaller proportion here than in any other borough.

#### Brooklyn Data Summary

```{r brooklyn, echo=FALSE}
my_boro <- 'BROOKLYN'
local_data <- subset(clean_data, BORO==my_boro)

num_shootings <- nrow(local_data)

brooklyn_shootings <- num_shootings

percent_of_shootings <- round((num_shootings / num_shootings_all) * 100, digits=2)

female_victims <- subset(local_data, VIC_SEX=='F')
male_victims <- subset(local_data, VIC_SEX=='M')
num_female_victims <- nrow(female_victims)
num_male_victims <- nrow(male_victims)
percent_female <- round((num_female_victims / num_shootings) * 100, digits=2)
percent_male <- round((num_male_victims / num_shootings) * 100, digits=2)

white_victims <- subset(local_data, VIC_RACE=='WHITE')
num_white_victims <- nrow(white_victims)
percent_white <- round((num_white_victims / num_shootings) * 100, digits=2)

black_victims <- subset(local_data, (VIC_RACE=='BLACK' | VIC_RACE=='BLACK HISPANIC'))
num_black_victims <- nrow(black_victims)
percent_black <- round((num_black_victims / num_shootings) * 100, digits=2)

hispanic_victims <- subset(local_data, (VIC_RACE=='WHITE HISPANIC' | VIC_RACE=='BLACK HISPANIC'))
num_hispanic_victims <- nrow(hispanic_victims)
percent_hispanic <- round((num_hispanic_victims / num_shootings) * 100, digits=2)

other_victims <- subset(local_data, (VIC_RACE!='WHITE' & VIC_RACE!='BLACK' & VIC_RACE!='WHITE HISPANIC' & VIC_RACE!='BLACK HISPANIC'))
num_other_victims <- nrow(other_victims)
percent_other <- round((num_other_victims / num_shootings) * 100, digits=2)

brook_shootings <- num_shootings
brook_fem <- percent_female
brook_mal <- percent_male
brook_white <- percent_white
brook_black<-percent_black
brook_hisp <- percent_hispanic
brook_oth <- other_victims
```

There were `r num_shootings` shooting incidents recorded by the NYPD in this data set in Brooklyn, `r percent_of_shootings`% of the total shooting incidents in New York City.

```{r brooklyn_shootings_graph, echo=FALSE}
pie(c(num_shootings, num_shootings_all-num_shootings), c("Shootings in Brooklyn", "Shootings Elsewhere in NYC"), main="Proportion of Shootings in NYC that occured in Brooklyn")
```

The victims of these shootings were `r percent_female`% female, and `r percent_male`% male.

```{r brooklyn_gender_graph, echo=FALSE}
vics <- c(num_female_victims, num_male_victims)
gender_labels <- c("Female Victims", "Male Victims")
pie(vics, gender_labels, main="Proportion of Victims by Gender")
```

`r percent_black`% of victims were Black and `r percent_hispanic`% were Hispanic, with only `r percent_white`% being White and non-Hispanic. `r percent_other`% of shooting victims belonged to other races or were listed as being of unknown race. Note that percentages sum to more than 100 because some victims are listed as both Black and Hispanic or White and Hispanic.

```{r brooklyn_race_graph, echo=FALSE}
races <- c(percent_black, percent_hispanic, percent_white, percent_other)
labels <- c("Black", "Hispanic", "White", "Other")
barplot(races, names.arg=labels, main="Shootings by Race", xlab="Race", ylab="Percentage of Shooting Victims")
```

##### Analysis

The plurality of shooting incidents in this data set occurred in Brooklyn, potentially making it the most dangerous borough. The gender and racial make up of shooting victims does not vary by a large margin from the city over all. Black victims are higher here, and Hispanic victims somewhat lower, but White victims and Female victims are still a very small percentage.

#### Manhattan Data Summary

```{r manhattan, echo=FALSE}
my_boro <- 'MANHATTAN'
local_data <- subset(clean_data, BORO==my_boro)

num_shootings <- nrow(local_data)

manhattan_shootings <- num_shootings

percent_of_shootings <- round((num_shootings / num_shootings_all) * 100, digits=2)

female_victims <- subset(local_data, VIC_SEX=='F')
male_victims <- subset(local_data, VIC_SEX=='M')
num_female_victims <- nrow(female_victims)
num_male_victims <- nrow(male_victims)
percent_female <- round((num_female_victims / num_shootings) * 100, digits=2)
percent_male <- round((num_male_victims / num_shootings) * 100, digits=2)

white_victims <- subset(local_data, VIC_RACE=='WHITE')
num_white_victims <- nrow(white_victims)
percent_white <- round((num_white_victims / num_shootings) * 100, digits=2)

black_victims <- subset(local_data, (VIC_RACE=='BLACK' | VIC_RACE=='BLACK HISPANIC'))
num_black_victims <- nrow(black_victims)
percent_black <- round((num_black_victims / num_shootings) * 100, digits=2)

hispanic_victims <- subset(local_data, (VIC_RACE=='WHITE HISPANIC' | VIC_RACE=='BLACK HISPANIC'))
num_hispanic_victims <- nrow(hispanic_victims)
percent_hispanic <- round((num_hispanic_victims / num_shootings) * 100, digits=2)

other_victims <- subset(local_data, (VIC_RACE!='WHITE' & VIC_RACE!='BLACK' & VIC_RACE!='WHITE HISPANIC' & VIC_RACE!='BLACK HISPANIC'))
num_other_victims <- nrow(other_victims)
percent_other <- round((num_other_victims / num_shootings) * 100, digits=2)

man_shootings <- num_shootings
man_fem <- percent_female
man_mal <- percent_male
man_white <- percent_white
man_black<-percent_black
man_hisp <- percent_hispanic
man_oth <- other_victims
```

There were `r num_shootings` shooting incidents recorded by the NYPD in this data set in Manhattan, `r percent_of_shootings`% of the total shooting incidents in New York City.

```{r manhattan_shootings_graph, echo=FALSE}
pie(c(num_shootings, num_shootings_all-num_shootings), c("Shootings in Manhattan", "Shootings Elsewhere in NYC"), main="Proportion of Shootings in NYC that occured in Manhattan")
```

The victims of these shootings were `r percent_female`% female, and `r percent_male`% male.

```{r manhattan_gender_graph, echo=FALSE}
vics <- c(num_female_victims, num_male_victims)
gender_labels <- c("Female Victims", "Male Victims")
pie(vics, gender_labels, main="Proportion of Victims by Gender")
```

`r percent_black`% of victims were Black and `r percent_hispanic`% were Hispanic, with only `r percent_white`% being White and non-Hispanic. `r percent_other`% of shooting victims belonged to other races or were listed as being of unknown race. Note that percentages sum to more than 100 because some victims are listed as both Black and Hispanic or White and Hispanic.

```{r manhattan_race_graph, echo=FALSE}
races <- c(percent_black, percent_hispanic, percent_white, percent_other)
labels <- c("Black", "Hispanic", "White", "Other")
barplot(races, names.arg=labels, main="Shootings by Race", xlab="Race", ylab="Percentage of Shooting Victims")
```

##### Analysis
Here, again, numbers do not vary much from city-wide proportions. Manhattan is perhaps the closest to the overall numbers for New York City.

#### Queens Data Summary

```{r queens, echo=FALSE}
my_boro <- 'QUEENS'
local_data <- subset(clean_data, BORO==my_boro)

num_shootings <- nrow(local_data)

queens_shootings <- num_shootings

percent_of_shootings <- round((num_shootings / num_shootings_all) * 100, digits=2)

female_victims <- subset(local_data, VIC_SEX=='F')
male_victims <- subset(local_data, VIC_SEX=='M')
num_female_victims <- nrow(female_victims)
num_male_victims <- nrow(male_victims)
percent_female <- round((num_female_victims / num_shootings) * 100, digits=2)
percent_male <- round((num_male_victims / num_shootings) * 100, digits=2)

white_victims <- subset(local_data, VIC_RACE=='WHITE')
num_white_victims <- nrow(white_victims)
percent_white <- round((num_white_victims / num_shootings) * 100, digits=2)

black_victims <- subset(local_data, (VIC_RACE=='BLACK' | VIC_RACE=='BLACK HISPANIC'))
num_black_victims <- nrow(black_victims)
percent_black <- round((num_black_victims / num_shootings) * 100, digits=2)

hispanic_victims <- subset(local_data, (VIC_RACE=='WHITE HISPANIC' | VIC_RACE=='BLACK HISPANIC'))
num_hispanic_victims <- nrow(hispanic_victims)
percent_hispanic <- round((num_hispanic_victims / num_shootings) * 100, digits=2)

other_victims <- subset(local_data, (VIC_RACE!='WHITE' & VIC_RACE!='BLACK' & VIC_RACE!='WHITE HISPANIC' & VIC_RACE!='BLACK HISPANIC'))
num_other_victims <- nrow(other_victims)
percent_other <- round((num_other_victims / num_shootings) * 100, digits=2)

queens_shootings <- num_shootings
queens_fem <- percent_female
queens_mal <- percent_male
queens_white <- percent_white
queens_black<-percent_black
queens_hisp <- percent_hispanic
queens_oth <- other_victims
```

There were `r num_shootings` shooting incidents recorded by the NYPD in this data set in the Queens, `r percent_of_shootings`% of the total shooting incidents in New York City.

```{r queens_shootings_graph, echo=FALSE}
pie(c(num_shootings, num_shootings_all-num_shootings), c("Shootings in Queens", "Shootings Elsewhere in NYC"), main="Proportion of Shootings in NYC that occured in Queens")
```

The victims of these shootings were `r percent_female`% female, and `r percent_male`% male.

```{r queens_gender_graph, echo=FALSE}
vics <- c(num_female_victims, num_male_victims)
gender_labels <- c("Female Victims", "Male Victims")
pie(vics, gender_labels, main="Proportion of Victims by Gender")
```

`r percent_black`% of victims were Black and `r percent_hispanic`% were Hispanic, with only `r percent_white`% being White and non-Hispanic. `r percent_other`% of shooting victims belonged to other races or were listed as being of unknown race. Note that percentages sum to more than 100 because some victims are listed as both Black and Hispanic or White and Hispanic.

```{r queens_race_graph, echo=FALSE}
races <- c(percent_black, percent_hispanic, percent_white, percent_other)
labels <- c("Black", "Hispanic", "White", "Other")
barplot(height=races, names.arg=labels, main="Shootings by Race", xlab="Race", ylab="Percentage of Shooting Victims", beside=TRUE, legend="")
```

##### Analysis
Continuing our pattern of robust similarity, the proportions in Queens do not deviate remarkably from those of the city overall. However, one statistic that does stand out is the number of shooting victims in our Other Races category. At 3.96% Queens has more than double the number of shooting victims in this category than New York City over all.

#### Staten Island Data Summary

```{r staten_island, echo=FALSE}
my_boro <- 'STATEN ISLAND'
local_data <- subset(clean_data, BORO==my_boro)

num_shootings <- nrow(local_data)

staten_island_shootings <- num_shootings

percent_of_shootings <- round((num_shootings / num_shootings_all) * 100, digits=2)

female_victims <- subset(local_data, VIC_SEX=='F')
male_victims <- subset(local_data, VIC_SEX=='M')
num_female_victims <- nrow(female_victims)
num_male_victims <- nrow(male_victims)
percent_female <- round((num_female_victims / num_shootings) * 100, digits=2)
percent_male <- round((num_male_victims / num_shootings) * 100, digits=2)

white_victims <- subset(local_data, VIC_RACE=='WHITE')
num_white_victims <- nrow(white_victims)
percent_white <- round((num_white_victims / num_shootings) * 100, digits=2)

black_victims <- subset(local_data, (VIC_RACE=='BLACK' | VIC_RACE=='BLACK HISPANIC'))
num_black_victims <- nrow(black_victims)
percent_black <- round((num_black_victims / num_shootings) * 100, digits=2)

hispanic_victims <- subset(local_data, (VIC_RACE=='WHITE HISPANIC' | VIC_RACE=='BLACK HISPANIC'))
num_hispanic_victims <- nrow(hispanic_victims)
percent_hispanic <- round((num_hispanic_victims / num_shootings) * 100, digits=2)

other_victims <- subset(local_data, (VIC_RACE!='WHITE' & VIC_RACE!='BLACK' & VIC_RACE!='WHITE HISPANIC' & VIC_RACE!='BLACK HISPANIC'))
num_other_victims <- nrow(other_victims)
percent_other <- round((num_other_victims / num_shootings) * 100, digits=2)

staten_shootings <- num_shootings
staten_fem <- percent_female
staten_mal <- percent_male
staten_white <- percent_white
staten_black<-percent_black
staten_hisp <- percent_hispanic
staten_oth <- other_victims
```

There were `r num_shootings` shooting incidents recorded by the NYPD in this data set in Staten Island, `r percent_of_shootings`% of the total shooting incidents in New York City.

```{r staten_island_shootings_graph, echo=FALSE}
pie(c(num_shootings, num_shootings_all-num_shootings), c("Shootings in Staten Island", "Shootings Elsewhere in NYC"), main="Proportion of Shootings in NYC that occured in Staten Island")
```

The victims of these shootings were `r percent_female`% female, and `r percent_male`% male.

```{r staten_island_gender_graph, echo=FALSE}
vics <- c(num_female_victims, num_male_victims)
gender_labels <- c("Female Victims", "Male Victims")
pie(vics, gender_labels, main="Proportion of Victims by Gender")
```

`r percent_black`% of victims were Black and `r percent_hispanic`% were Hispanic, with only `r percent_white`% being White and non-Hispanic. `r percent_other`% of shooting victims belonged to other races or were listed as being of unknown race. Note that percentages sum to more than 100 because some victims are listed as both Black and Hispanic or White and Hispanic.

```{r staten_island_race_graph, echo=FALSE}
races <- c(percent_black, percent_hispanic, percent_white, percent_other)
labels <- c("Black", "Hispanic", "White", "Other")
barplot(races, names.arg=labels, main="Shootings by Race", xlab="Race", ylab="Percentage of Shooting Victims")
```

##### Analysis
Staten Island had the fewest shooting incidents recorded in this sample. A couple of numbers jump out here. First, the proportion of White victims is much higher here than elsewhere in the city. Almost 4 times higher than the city-wide average. The number of female victims is also higher here by a couple percentage points. But still, Black victims are the majority by a huge margin. Victims are also still overwhelmingly male.

### Conclusions

```{r conc_shootings_pie, echo=FALSE}
props <- c(bronx_shootings, brooklyn_shootings, manhattan_shootings, queens_shootings, staten_island_shootings)
labels <- c("Bronx", "Brooklyn", "Manhattan", "Queens", "Staten Island")
pie(props, labels, main="Shootings in NYC by Borough")
```

```{r conc_shootings_bar, echo=FALSE}
fem <- c(Bronx=bronx_fem, Brooklyn=brook_fem, Manhattan=man_fem, Queens=queens_fem, Staten_Island=staten_fem)
mal <- c(Bronx=bronx_mal, Brooklyn=brook_mal, Manhattan=man_mal, Queens=queens_mal, Staten_Island=staten_mal)

white <- c(Bronx=bronx_white, Brooklyn=brook_white, Manhattan=man_white, Queens=queens_white, Staten_Island=staten_white)
black <- c(Bronx=bronx_black, Brooklyn=brook_black, Manhattan=man_black, Queens=queens_black, Staten_Island=staten_black)
hisp <- c(Bronx=bronx_hisp, Brooklyn=brook_hisp, Manhattan=man_hisp, Queens=queens_hisp, Staten_Island=staten_hisp)
oth <- c(Bronx=bronx_oth, Brooklyn=brook_oth, Manhattan=man_oth, Queens=queens_oth, Staten_Island=staten_oth)

race_legend_labels <- c("Black", "Hispanic", "White")
gender_legend_labels <- c("Male", "Female")

gender_data <- rbind(mal, fem)
race_data <- rbind(black, hisp, white)

barplot(gender_data, main="Shootings by Gender", xlab="Gender", ylab="Percentage of Shooting Victims", beside=TRUE, legend=gender_legend_labels) 

barplot(race_data, main="Shootings by Race", xlab="Race", ylab="Percentage of Shooting Victims",beside=TRUE, legend=race_legend_labels)


```


As you can see from the pie chart above, the number shooting incident numbers vary significantly from borough to borough. There was, to me at least, a remarkable level of similarity in racial and gender make up of victims though. Nowhere in New York City are Black victims less than 70% of shooting victims and males make up around 90% of shooting victims everywhere in the city. Gender proportions in particular are remarkably stable regardless of borough. 

One interesting correlation that emerges is that boroughs with fewer shooting incidents seem to have a higher proportion of white victims.



##### Further Research Suggestions
Further research is necessary to determine what effects if any, may be confounding this data. Of particular concern would be the gender and racial make up of the general population of each borough. While it currently appear to be the case that boroughs with fewer incidents have a greater proportion of White victims, it may only be the case that those boroughs have a higher proportion of White people in the general population. It may also be that those boroughs have smaller populations over all, suggesting the possibility that it is Whiter populations or smaller populations, respectively, which are causing the higher proportion of White victims.

It is also worth researching the particular causes for the extremely high proportion of shooting victims who are identified as Black or Hispanic. These proportions seem to far exceed the expected proportion given their proportions in the general population.

#### Possible Biases


### Session Info
```{r session_info, echo=FALSE}
sessionInfo()
```

