---
title: "NYPD Precinct 52 Shooting Incident Victim Data"
author: Jacob Moore
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data Source
[Webpage](https://catalog.data.gov/dataset/nypd-shooting-incident-data-historic/resource/c564b578-fd8a-4005-8365-34150d306cc4)

[CSV](https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD)

```{r define_shooting_data, echo=FALSE}
shooting_data <- read.csv("https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD")
num_shootings_all <- nrow(shooting_data)
```

```{r clean_data, echo=FALSE}

clean_data_columns <- subset(shooting_data, select = -c(INCIDENT_KEY, OCCUR_DATE, OCCUR_TIME, BORO, PERP_AGE_GROUP, PERP_SEX, PERP_RACE, JURISDICTION_CODE, LOCATION_DESC, STATISTICAL_MURDER_FLAG, X_COORD_CD, Y_COORD_CD, Latitude, Longitude, Lon_Lat))

clean_data_rows <- subset(clean_data_columns, PRECINCT==52)

clean_data <- subset(clean_data_rows, select=-c(PRECINCT))
```

### Data Summary
```{r data_summary, echo=FALSE}
num_shootings <- nrow(clean_data)

percent_52 <- round((num_shootings / num_shootings_all) * 100, digits=2)

female_victims <- subset(clean_data, VIC_SEX=='F')
male_victims <- subset(clean_data, VIC_SEX=='M')
num_female_victims <- nrow(female_victims)
num_male_victims <- nrow(male_victims)
percent_female <- round((num_female_victims / num_shootings) * 100, digits=2)
percent_male <- round((num_male_victims / num_shootings) * 100, digits=2)

white_victims <- subset(clean_data, VIC_RACE=='WHITE')
num_white_victims <- nrow(white_victims)
percent_white <- round((num_white_victims / num_shootings) * 100, digits=2)

black_victims <- subset(clean_data, (VIC_RACE=='BLACK' | VIC_RACE=='BLACK HISPANIC'))
num_black_victims <- nrow(black_victims)
percent_black <- round((num_black_victims / num_shootings) * 100, digits=2)

hispanic_victims <- subset(clean_data, (VIC_RACE=='WHITE HISPANIC' | VIC_RACE=='BLACK HISPANIC'))
num_hispanic_victims <- nrow(hispanic_victims)
percent_hispanic <- round((num_hispanic_victims / num_shootings) * 100, digits=2)

other_victims <- subset(clean_data, (VIC_RACE!='WHITE' & VIC_RACE!='BLACK' & VIC_RACE!='WHITE HISPANIC' & VIC_RACE!='BLACK HISPANIC'))
num_other_victims <- nrow(other_victims)
percent_other <- round((num_other_victims / num_shootings) * 100, digits=2)
```

There were `r num_shootings_all` shooting incidents involving the NYPD. `r num_shootings` of those occurred in Precinct 52, accounting for `r percent_52`% of overall shootings. 

The victims of these shootings were `r percent_female`% female, and `r percent_male`% were male.

Racially, `r percent_black`% of victims in Precinct 52 were Black and `r percent_hispanic`% were Hispanic, with only `r percent_white`% being White and non-Hispanic. `r percent_other`% of shooting victims belonged to other races or were listed as being of unknown race. Note that percentages sum to more than 100 because some victims are listed as both Black and Hispanic or White and Hispanic.

Victims of shootings involving Precinct 52 were thus, overwhelmingly male and non-white.
